\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual Multi-Loop Integrands

### 2023-09-22 notes/2loopNLSM.tex
fixed BEL appendix

It is instructive to see how both of these constructions produce valid tree-level amplitudes for the pion. First we'll start with two $Y$ particles on legs 1 and 4. Applying the Feynman rules above, and plugging in on-shell states for the $Z$-particles produces the following $s$- and $t$-channel numerators:
\begin{align}
n^{YY}_s &= (T^2T^3)_{14} = s_{12}^2 
\\
 n^{YY}_t &=  F^{3}_{\,2|X}T^X_{14}  = s_{14}(s_{13}-s_{12})
\end{align}
Plugging these numerators into the ordered amplitudes $A(s,t)$ yields the desired result:
\be\label{eq:NLSMYZ4point}
A^{YY}_{(s,t)} = \frac{n^{YY}_s}{s_{12}}+\frac{ n^{YY}_t }{s_{14}} = s_{13}
\end{equation}
Similarly we can do the same for the $Z$ and $\bar{Z}$ configuration. Below we have plugged in the on-shell $Z$-particle states, but have left the $\bar{Z}$ index free. This produces the following numerators:
\begin{align}
n^{\bar{Z}Z}_s &= {}_4(F^{3}F^{2})_{1} =  s_{12}(s_{14}-s_{13})p_2^{\mu_1}+s_{12}^2(p_3-p_4)^{\mu_1}
\\
 n^{\bar{Z}Z}_t &=   {}_2(F^{3}F^{4})_{1}  = s_{14}(s_{12}-s_{13})p_4^{\mu_1}+s_{14}^2(p_3-p_2)^{\mu_1}
\end{align}
where we have defined the short hand notation:
\begin{equation}
{}_x(F^{a_1}F^{a_2}\cdots F^{a_n})_{y} \equiv F^{a_1}_{\,x|b_2}F^{a_2}_{\,b_2|b_3}\cdots F^{a_n}_{\,b_n|y}
\end{equation}
Similarly, we find the following ordered amplitude when plugging in these numerators:
\begin{equation}
A^{\bar{Z}Z}_{(s,t)} = \frac{n^{\bar{Z}Z}_s}{s_{12}}+\frac{ n^{\bar{Z}Z}_t }{s_{14}} = -s_{13}(p_2+p_3+p_4)^{\mu_1} = s_{13} \,p_1^{\mu_1}
\end{equation}
Plugging in the on-shell polarizatoin of the conjugate field in \cref{eq:onShellZStates} produces precisely the desired result of \cref{eq:NLSMYZ4point}. Indeed, this construction is valid to all orders at tree-level. One can see this by considering the only two possible factorization channels the contribute the each of these amplitudes, the $YY$ cut and the $\bar{Z}Z$ cut:
\begin{align}
A(...,Y,...,Y,...) &\rightarrow A(...,Y,...,Y)A(Y,...,Y,...)
\\
&\rightarrow A(...,Y,...,Y,...,Z)A(\bar{Z},...)
\end{align}
Both factorization channels are valid descriptions of NLSM amplitudes when plugging in the $Z$ and $\bar{Z}$ on-shell states. As constructed, the $Y$ and $\bar{Z}$ particles may be off-shell, while the $Z$ particles must be placed on shell, $\varepsilon^Z_\mu(k)\rightarrow k_\mu$ in order for these numerators to produce NLSM ordered amplitudes. Given this special property of $YZ$-theory (namely that the $Z\bar{Z}$ and $YY$ constrctions are equivalent), the $Y$ particle need not be considered when constructing NLSM amplitudes on shell. This is essentially what $J$-theory does, as we will show below.

@@

N^{n\text{-gon},YZ}_{(12...n)} = \langle T^{1}T^{2}\cdots T^{n}\rangle+ \rangle F^1F^2\cdots F^n\rangle

@@

where $(\,\cdots)$ indicates an internal contraction over the $YY$ and

@@

N_{(12...n)}^{\text{NLSM}}=(T^{1}T^{2}\cdots T^{n}) = (\ell_1 k_1)(\ell_2 k_2) \cdots (\ell_n k_n)

@@

and similarly so for the internal vector contribution:

@@

 (F^1F^2\cdots F^n) = D (\ell_1 k_1)(\ell_2 k_2) \cdots (\ell_n k_n) + \mathcal{O}(D^0)

@@

desired pion cuts. In order for the Feynman rules for $Y\!Z$ theory
compute one-loop color-dual numerators consistent with NLSM cuts, we
must add some additional states to cancel off he spurious poles, while

@@

While the $\bar{Z}Z$ vector loop spoils color-kinematics off-shell, the $Y\!Z$ loop alone gives us a desired expression for the $n$-gon. The important takeaway is that we now have a guess for the form of the off-shell three-point vertex that has a chance of manifesting the duality off-shell. The $n$-gon numerator above has scalar insertions of the following kinematic vertex:
\begin{equation}
T^{a}_{LR} = k_a(\ell_L-\ell_{R}) = (\ell_L+\ell_{R}) (\ell_L-\ell_{R})  = \ell_L^2-\ell_{R}^2 
\end{equation}
Given this structure, in the next section we will attempt to construct two-loop basis diagrams from these cubic vertex assignments and try to reverse engineer the particle content that produces these master numerators.

Before proceeding, we note a strange property of the $n$-gon numerator for the pions. In this form, Jacobi relations produce \textit{non-vanishing} values for bubbles on external legs (BELs). However, it is easy to see that the contribution integrates to zero after applying IBP relations/tensor reduction on the bubble. The BEL diagram can be reconstructed from the $n$-gon as follows:
\begin{equation}
N^{\text{BEL}}_{1|2,34} = N^{\text{NLSM}}_{(1234,\ell)}-N^{\text{NLSM}}_{(1243,\ell)}-N^{\text{NLSM}}_{(1342,\ell)}+N^{\text{NLSM}}_{(1432,\ell)}
\end{equation}
where we define the loop momentum to be in between the left most and right most leg on the box. Plugging in particular values for $\ell_i$, we obtain the following expression for the BEL
\begin{equation}
N^{\text{BEL}}_{1|2,34} = s_{12} (\ell+k_1)^2 \ell^{\mu} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} 
\end{equation}
Notice there is an overall factor that cancels one of the propagators. Plugging this in, produces an integral of the following form
\begin{equation}
\mathcal{I}^{\text{BEL}}_{1|2,34} = s_{12} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} \int \frac{d^D \ell}{i\pi^{D/2}} \frac{\ell^\mu }{\ell^2-\mu^2} \sim   s_{12}(s_{13}-s_{14}) (\mu^2)^{D/2}
\end{equation}
where we have introduced a mass regulator that will be proportional the the on-shell momentum inside the BEL, $\mu^2 \equiv k_1^2$. Thus, in sufficiently large dimension, $D>2$, this integral suppresses the $\mu^{-2}$ divergence appearing in the denominator of the BEL diagram.

>>>>

It is instructive to see how both of these constructions produce valid tree-level amplitudes for the pion. First we'll start with two $Y$ particles on legs 1 and 4. Applying the Feynman rules above, and plugging in on-shell states for the $Z$-particles produces the following $s$- and $t$-channel numerators:
\begin{align}
n^{YY}_s &= (T^2T^3)_{14} = s_{12}^2 
\\
 n^{YY}_t &=  F^{3}_{\,2|X}T^X_{14}  = s_{14}(s_{13}-s_{12})
\end{align}
Plugging these numerators into the ordered amplitudes $A(s,t)$ yields the desired result:
\be\label{eq:NLSMYZ4point}
A^{YY}_{(s,t)} = \frac{n^{YY}_s}{s_{12}}+\frac{ n^{YY}_t }{s_{14}} = s_{13}
\end{equation}
Similarly we can do the same for the $Z$ and $\bar{Z}$ configuration. Below we have plugged in the on-shell $Z$-particle states, but have left the $\bar{Z}$ index free. This produces the following numerators:
\begin{align}
n^{\bar{Z}Z}_s &= {}_4(F^{3}F^{2})_{1} =  s_{12}(s_{14}-s_{13})p_2^{\mu_1}+s_{12}^2(p_3-p_4)^{\mu_1}
\\
 n^{\bar{Z}Z}_t &=   {}_2(F^{3}F^{4})_{1}  = s_{14}(s_{12}-s_{13})p_4^{\mu_1}+s_{14}^2(p_3-p_2)^{\mu_1}
\end{align}
where we have defined the short hand notation:
\begin{equation}
{}_x(F^{a_1}F^{a_2}\cdots F^{a_n})_{y} \equiv F^{a_1}_{\,x|b_2}F^{a_2}_{\,b_2|b_3}\cdots F^{a_n}_{\,b_n|y}
\end{equation}
Similarly, we find the following ordered amplitude when plugging in these numerators:
\begin{equation}
A^{\bar{Z}Z}_{(s,t)} = \frac{n^{\bar{Z}Z}_s}{s_{12}}+\frac{ n^{\bar{Z}Z}_t }{s_{14}} = -s_{13}(p_2+p_3+p_4)^{\mu_1} = s_{13} \,p_1^{\mu_1}
\end{equation}
Plugging in the on-shell polarizatoin of the conjugate field in \cref{eq:onShellZStates} produces precisely the desired result of \cref{eq:NLSMYZ4point}. Indeed, this construction is valid to all orders at tree-level. One can see this by considering the only two possible factorization channels the contribute the each of these amplitudes, the $YY$ cut and the $\bar{Z}Z$ cut:
\begin{align} \label{YZCut1}
A(...,Y,...,Y,...) &\rightarrow A(...,Y,...,Y)A(Y,...,Y,...)
\\
&\rightarrow A(...,Y,...,Y,...,Z)A(\bar{Z},...) \label{YZCut2}
\end{align}
Both factorization channels are valid descriptions of NLSM amplitudes when plugging in the $Z$ and $\bar{Z}$ on-shell states. As constructed, the $Y$ and $\bar{Z}$ particles may be off-shell, while the $Z$ particles must be placed on shell, $\varepsilon^Z_\mu(k)\rightarrow k_\mu$ in order for these numerators to produce NLSM ordered amplitudes. Given this special property of $YZ$-theory (namely that the $Z\bar{Z}$ and $YY$ constrctions are equivalent), the $Y$ particle need not be considered when constructing NLSM amplitudes on shell. This is essentially what $J$-theory does, as we will show below.

@@

N^{n\text{-gon},YZ}_{(12...n)} = \langle T^{1}T^{2}\cdots T^{n}\rangle+ \langle F^1F^2\cdots F^n\rangle

@@

where $\langle\,\cdots\rangle$ indicates an internal contraction over the $YY$ and

@@

N_{(12...n)}^{\text{NLSM}}=\langle T^{1}T^{2}\cdots T^{n}\rangle =2^n (\ell_1 k_1)(\ell_2 k_2) \cdots (\ell_n k_n)

@@

which is what we found in the text. Similarly, there is a pure vector contribution that is identical to the semi-abelian YM $n$-gon numerator that we constructed in \cref{SDYMnGon} after projecting external states along longitudinal modes $\varepsilon \rightarrow k$:

@@

 \langle F^1F^2\cdots F^n \rangle= D (\ell_1 k_1)(\ell_2 k_2) \cdots (\ell_n k_n) + \mathcal{O}(D^0)

@@

desired pion cuts due to the factorization of \cref{YZCut1} and \cref{YZCut2}. In order for the Feynman rules for $Y\!Z$ theory
compute one-loop color-dual numerators consistent with NLSM cuts, one would have to add some additional states to cancel off he spurious poles, while

@@

While the $\bar{Z}Z$ vector loop spoils color-kinematics off-shell, the $\bar{Y}Y$ loop alone gives us a desired expression for the $n$-gon. However, as we noted in the text, the $n$-gon numerator have the strange property that they produce \textit{non-vanishing} values for bubbles on external legs (BELs). However through direct integration of the BEL integrand, it is easy to see that the contribution integrates to zero after applying IBP relations/tensor reduction on the bubble. The BEL diagram can be reconstructed from the $n$-gon as follows:
\begin{equation}
N^{\text{BEL}}_{1|2,34} = N^{\text{NLSM}}_{(1234|\ell)}-N^{\text{NLSM}}_{(1243|\ell)}-N^{\text{NLSM}}_{(1342|\ell)}+N^{\text{NLSM}}_{(1432|\ell)}
\end{equation}
where we define the loop momentum to be in between the left most and right most leg on the box. Plugging in particular values for $\ell_i$, we obtain the following expression for the BEL:
\begin{equation}
N^{\text{BEL}}_{1|2,34} = s_{12} (\ell+k_1)^2 \ell^{\mu} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} \,.
\end{equation}
Notice there is an overall factor that cancels one of the propagators. Plugging this in, produces an integral of the following form
\begin{equation}
\mathcal{I}^{\text{BEL}}_{1|2,34} = s_{12} k_1^{\nu} k_2^{[\mu} k_{[34]}^{\nu]} \int \frac{d^D \ell}{i\pi^{D/2}} \frac{\ell^\mu }{\ell^2-\mu^2} \sim   s_{12}(s_{13}-s_{14}) (\mu^2)^{D/2}
\end{equation}
where we have introduced a mass regulator that will be proportional the the on-shell momentum inside the BEL, $\mu^2 \equiv k_1^2$. Thus, in sufficiently large dimension, $D>2$, this integral suppresses the $\mu^{-2}$ divergence appearing in the denominator of the BEL diagram.
